---
title: "Reproducible Research"
output:
  html_document:
    df_print: paged
---
## Data Wrangling:
```{r}
windowsFont("Lucida Console")
```

### Session Information
```{r Session Info}
sessionInfo()
```

### Loading Essential Packages
```{r Loading Packages}
library(dplyr)
library(ggplot2)
```

### The Dataset
```{r Loading the dataset}
activity = read.csv("activity.csv")
```
Here are the summary statistics of the dataset
```{r General Chracteristics, warning FALSE}
summary(activity)
```

## Exploratory Data Analysis
### Numerical Summaries
```{r Grouping, warning = FALSE}
## Mean and Median number of steps taken each day by grouping daywise
steps_details <- activity %>%
  group_by(date) %>%
  summarise(steps_per_day = sum(steps, na.rm = FALSE), median_per_day = median(steps),
            mean_per_day = mean(steps))
head(steps_details)
```

### Plots
```{r Histogram,warning=FALSE}
## Plotting a Histogram
ggplot(mapping = aes(x = steps_details$steps_per_day))+
  geom_histogram()+ 
  ggtitle("Histogram of total steps per day")+
  xlab("Total Steps per Day") 
  
```

```{r Time Series Plot, warning=FALSE}
## Average steps during each time interval
# Grouping the variables on the basis of the interval
interval_details <- activity %>%
  group_by(interval) %>%
  summarise(mean_steps = mean(steps, na.rm = TRUE))

# Time series plot
ggplot(interval_details, mapping = aes(x=interval, y=mean_steps))+
  geom_line()+
  ggtitle("Time Series Plot of Steps Averaged over all days")+
  ylab("Steps Averaged over all days")
```

Calculating the interval during which mean number of steps is the highest: 
```{r MaxInterval, warning = FALSE}
interval_details$interval[which.max(interval_details$mean_steps)]
```

## Data Cleaning 

We first get a count of the missing values:
```{r NA counts, warning = FALSE}
lapply(activity, function(x) sum(is.na(x)))
```

### Imputing the missing Values:
Only 'steps' has missing values. Hence, replace them by the mean and store the analytic data in a new dataset - 'activity_ana'

```{r Imputing Missing Values}
activity_ana <- activity
activity_ana$steps[is.na(activity_ana$steps)] <- 
  mean(activity_ana$steps, na.rm = TRUE)
```

### Plots
```{r Histogram without NA, warning=FALSE}
## Mean and Median number of steps taken each day by grouping daywise
activity_ana%>%
  group_by(date) %>%
  summarise(steps_per_day = sum(steps)) %>%
  
  ggplot(steps_per_day, mapping = aes(x = steps_details$steps_per_day))+
  geom_histogram(binwidth = 1000)+
  ggtitle("Histogram of total steps per day")+
  xlab("Total Steps per Day") 
```

## Are there differences in activity patterns between weekdays and weekends?
```{r Factor Variables, warning = FALSE}
# Creating the factor variable 'day_type'
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
activity_ana$day_type <- 
  factor((weekdays(as.Date(activity_ana$date)) %in% weekdays1), 
         levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))

# Boxplot to compare the steps during weekdays and weekends
ggplot(activity_ana, mapping = aes(x = day_type, y = log10(steps)))+
    geom_boxplot() +
    theme(axis.title.x = element_blank())
```

**Narrative:**
- The number of steps is less variable during weekends
- The median number of steps is slightly more during weekends
